Dependent Object Types

### 2012-07-01 doc/semester/report.tex
formatting, typos

The DOT calculus is expressed using a small-step operational
semantics. The typing judgement directly or indirectly uses the
following other judgements: membership, expansion, well-formedness and
subtyping. Expansion ``flattens'' all the declarations of a type, and
membership selects a particular flattened declaration by
label. Expansion is precise. The \textsc{new} typing judgement relies
on the preciseness of expansion to check that all bounds of types
declared within the type of the constructor are consistent (i.e. that
the lower bound is a subtype of the upper bound), and that all values
declared within the type of the constructor are initialized.

@@

There are two membership ($t \ni D$) rules: one for when the term $t$
is a path, and one for an arbitrary term $t$. For paths, we can
substitute the self-references in the declarations, but we cannot do
so for arbitrary terms as the resulting types wouldn't be well-formed
syntactically. Hence, the $\textsc{term-$\ni$}$ has the restriction
that self-ocurrences are not allowed. Here is a counterexample related
to this restriction.

@@

\begin{align*}
&\mlnew {x_0} {\Top \mlrefine z { \Ldecl {A_1} \Bot {\Top \mlrefine z {\\
&\gap\Ldecl {A_2} \Bot \Top\\
&\gap\Ldecl {A_3} \Bot \Top\\
&\gap\Ldecl {C_2} \Bot {z.A_2}}}}\ldefs{}}{
\mlnew {x_1} {\Top \refine z {\Ldecl {C_1} \Bot {\Top \refine z {\Ldecl {A_1} \Bot {x_0.A_1}}}}\ldefs{}}}{
\mlnew {x_2} {{x_1.C_1} \refine z {\Ldecl {A_1} \Bot {{x_0.A_1} \refine z {\Ldecl {A_2} \Bot {z.A_3}}}}\ldefs{}}}{
\mlnew {x_3} {{x_1.C_1} \refine z {\Ldecl {A_1} \Bot {{x_0.A_1} \refine z {\Ldecl {A_3} \Bot {z.A_2}}}}\ldefs{}}}{
\mlapp {\abs x {x_1.C_1} {(\abs {z_0} {x.A_1 \tand x_3.A_1} {\new z {z_0.C_2} {\app {(\abs x \Top x)} z}})}} {\ x_2}}
\end{align*}

@@

We introduce a new judgement form for whether a type is well-formed
and expanding: $\Gamma \ts T \wfe$ if and only if $\Gamma \ts T \wf$
and $\exists \seq D$ such that $\Gamma \ts T \expand_z \seq D$. Then,
we replace all other uses of the $\wf$ judgment, including those
within the $\wf$ judgment inference rules, with uses of the $\wfe$
judgment.

@@

\section{Conclucion and Future Work}\label{conclusion}
This semester's work was instructive in pointing out the problematic
aspects of the original DOT calculus, and exploring some possible
fixes. However, I haven't been able to reach a revised calculus
without soundness hole. Furthermore, I believe that the revised
calculus with all the patches is not as elegant as the original
calculus. Going forward, I want to take a broader view by by studying
and incorporating ideas from other calculi which models dependent
types, and path-dependent types in particular.

>>>>

The DOT calculus is expressed using a small-step operational
semantics. The typing judgment directly or indirectly uses the
following other judgments: membership, expansion, well-formedness and
subtyping. Expansion ``flattens'' all the declarations of a type, and
membership selects a particular flattened declaration by
label. Expansion is precise. The \textsc{new} typing judgment relies
on the preciseness of expansion to check that all bounds of types
declared within the type of the constructor are consistent (i.e. that
the lower bound is a subtype of the upper bound), and that all values
declared within the type of the constructor are initialized.

@@

There are two membership ($t \ni D$) rules: one for when the term $t$
is a path, and one for an arbitrary term $t$. For paths, we can
substitute the self-references in the declarations, but we cannot do
so for arbitrary terms as the resulting types wouldn't be well-formed
syntactically. Hence, the $\textsc{term-$\ni$}$ has the restriction
that self-occurrences are not allowed. Here is a counterexample related
to this restriction.

@@

\begin{align*}
&\mlnew {x_0} {\Top \mlrefine z { \Ldecl {A_1} \Bot {\Top \mlrefine z {\\
&\gap\Ldecl {A_2} \Bot \Top\\
&\gap\Ldecl {A_3} \Bot \Top\\
&\gap\Ldecl {C_2} \Bot {z.A_2}}}}\ldefs{}}{
\mlnew {x_1} {\Top \refine z {\Ldecl {C_1} \Bot {\Top \refine z {\Ldecl {A_1} \Bot {x_0.A_1}}}}\ldefs{}}}{
\mlnew {x_2} {{x_1.C_1} \refine z {\Ldecl {A_1} \Bot {{x_0.A_1} \refine z {\Ldecl {A_2} \Bot {z.A_3}}}}\ldefs{}}}{
\mlnew {x_3} {{x_1.C_1} \refine z {\Ldecl {A_1} \Bot {{x_0.A_1} \refine z {\Ldecl {A_3} \Bot {z.A_2}}}}\ldefs{}}}{
\mlapp {\abs x {x_1.C_1} {(\abs {z_0} {x.A_1 \tand x_3.A_1} {\\&\gap\gap\gap\new z {z_0.C_2} {\app {(\abs x \Top x)} z}})}} {\ x_2}}
\end{align*}

@@

We introduce a new judgment form for whether a type is well-formed
and expanding: $\Gamma \ts T \wfe$ if and only if $\Gamma \ts T \wf$
and $\exists \seq D$ such that $\Gamma \ts T \expand_z \seq D$. Then,
we replace all other uses of the $\wf$ judgment, including those
within the $\wf$ judgment inference rules, with uses of the $\wfe$
judgment.

@@

\section{Conclusion and Future Work}\label{conclusion}
This semester's work was instructive in pointing out the problematic
aspects of the original DOT calculus, and exploring some possible
fixes. However, I haven't been able to reach a revised calculus
without soundness hole. Furthermore, I believe that the revised
calculus with all the patches is not as elegant as the original
calculus. Going forward, I want to take a broader view by by studying
and incorporating ideas from other calculi which models dependent
types, and path-dependent types in particular.
